A global compilation reveals divergent biomass partitioning between plant functional types

### 2014-11-16 manuscript/draftpaper.Rnw
Removed old climate analysis, added new GAM-based general one. Added as new Table 1.

load("tables/Tables_varpart.RData")
baad <- readRDS("../data/baad.rds")
@

@@

\textbf{The partitioning of biomass in terrestrial vegetation between leaves, stems and roots is a critical component of dynamic global vegetation models (DGVMs). Biomass partioning is consistently shown to be one of the most uncertain and important drivers for global predictions of carbon storage \cite{ise_comparison_2010}, the lifetime of carbon in ecosystems \cite{friend_carbon_2014}, and the CO$_2$ response of net primary production \cite{dekauwe2014does} . Little is known about direct and indirect effects of climate on biomass partitioning, largely due to the lack of a global individual-based biomass dataset. Here, we analyze the largest database on plant biomass to date [cite]. We show that partitioning of biomass to leaves vs. woody stems is strongly dependent on the plant functional type. Evergreen gymnosperms maintain three times more leaf biomass at a given total aboveground biomass than deciduous angiosperms, with evergreen angiosperms taking an intermediate value. We found that leaf mass per area, an indicator of leaf construction cost, explains these differences so that the PFTs converged in the ratio  of leaf area to aboveground biomass. We found weak effects of mean annual temperature and no effects of precipitation on biomass partitioning, suggesting that climate-related differences are captured largely by a shift in PFT. These results point to a simple algorithm for biomass partitioning in the world's forests.}

@@

@@

<<tablepreload2>>=
library(car)
load("tables/Fits_lme_mlfmst_MAPMAT.RData")
printanova <- function(x, caption, label=NULL){
  rownames(x) <- gsub("log10(m.st)", "MS", rownames(x), fixed=TRUE) 
  rownames(x) <- gsub("pft", "PFT", rownames(x)) 
  print(xtable(x, caption=caption, label=label),
                                        caption.placement="top")
}
@

<<results='asis'>>=
printanova(Anova(mlfmst_lme1), 
           label="table:matlmeanova",
           caption="Type-II Anova table for a linear mixed-effects model of $M_F$ as a function of $M_S$ (both log-transformed), PFT and mean annual temperature (MAT). Species by dataset was used as the random effect. Further analysis of the interactions revealed that MAT only significantly affected the relationship between $M_F$ and $M_S$ for the deciduous angiosperms (not shown).")
@

<<results='asis'>>=
printanova(Anova(mlfmst_lme2), 
           label="table:maplmeanova",
           caption="Type-II Anova table for a linear mixed-effects model of $M_F$ as a function of $M_S$ (both log-transformed), PFT and mean annual precipitation (MAT). Species by dataset was used as the random effect. We did not analyze MAT and MAP together in one model, as they were strongly correlated, making interpretation of the fitted model difficult.")
@

>>>>

load("tables/Tables_varpart.RData")
load("tables/gamr2MIgdd0.RData")
baad <- readRDS("../data/baad.rds")
@

@@

\textbf{The partitioning of biomass in terrestrial vegetation between leaves, stems and roots is a critical component of dynamic global vegetation models (DGVMs). Biomass partioning is consistently shown to be one of the most uncertain and important drivers for global predictions of carbon storage \cite{ise_comparison_2010}, the lifetime of carbon in ecosystems \cite{friend_carbon_2014}, and the CO$_2$ response of net primary production \cite{dekauwe2014does} . Little is known about direct and indirect effects of climate on biomass partitioning, largely due to the lack of a global individual-based biomass dataset. Here, we analyze the largest database on plant biomass to date [cite]. We show that partitioning of biomass to leaves vs. woody stems is strongly dependent on the plant functional type. Evergreen gymnosperms maintain three times more leaf biomass at a given total aboveground biomass than deciduous angiosperms, with evergreen angiosperms taking an intermediate value. We found that leaf mass per area, an indicator of leaf construction cost, explains these differences so that the PFTs converged in the ratio  of leaf area to aboveground biomass. We found weak effects of climate while accounting for PFT effects, suggesting that climate-related differences are captured largely by a shift in PFT. These results can be used to constrain biomass partitioning routines in global dynamic vegetation models.}

@@

<<results='asis'>>=
print(xtable(g0, label="table/gamr2table",
      caption="Explained variance by plant functional type (PFT), plant height (H), moisture index (MI) and growing season average temperature (mgdd0), for six whole-plant variables."),
      include.rownames=FALSE, caption.placement="top")
@

@@
